Fail Faster

### 2025-03-24 paper/paper.tex
blah
eval para

We evaluate the optimizations with a series of case studies in each
generator library, and show that the \name{} generators run faster
than their unstaged equivalents.  Further, we show generators that run
faster also find bugs faster by running these case studies in the Etna
platform~\cite{etna}, which injects bugs into a system and measures
how quickly various generators detect them.  \jwc{Beef up this
  paragraph, explain the results of our eval.}  \todo{We precisely
  quantify the impact of each intervention and explain how they play
  distinct yet complementary roles in improving generator performance.
}

@@

Back in the ML-family world, the primary use of staging is building embedded DSLs
with minimal performance overhead, an idea which has come to be known as ``abstraction without regret''
\cite{lms,parreaux17,parreauxthesis, sheard99, trattdsl}.
This technique has been applied across a range of domains, including
Big data processing \cite{jet12}, stream processing \cite{moller20,strymonas},
query processing \cite{rhyme}, and parser combinators \cite{sspc, staged-parsers, krishnaswami19, flap}.
This work on staged parsers is the closest analogue to ours---indeed, PBT research has drawn an explicit link between parsing and generation, framing generators
as ``parsers of randomness'' \cite{goldsteinParsingRandomness2022}.
However, this equivalence is primarily theoretical. PBT generators are (a) not actually implemented as parsers
of random sequences, and (b) support choice-based combinators like \texttt{weighted\_union} that parsers do not.
For this reason, we must build staged generator libraries from scratch and cannot simply use existing staged parser libraries

>>>>

We evaluate the optimizations with a series of case studies in each
generator library, and show that the \name{} generators run faster
than their unstaged equivalents. In \camlname, they run up to $7\times$
faster, and in \scalaname, up to $13\times$ faster. Using our improved version
of \bq's \rand in \camlname, we see even greater gains---more than $12\times$---showing that
both staging and fast randomness play distinct and complementary roles in improving generator performance.
Further,
we show that generators that run
faster also find bugs faster by running our case studies in the Etna
platform~\cite{etna}, which injects bugs into a system and measures
how quickly various generators detect them. In Etna, we find bugs up 
to $3.8\times$ faster on average, and up to $13\times$ in some cases.

@@

Back in the ML-family world, the primary use of staging is building embedded DSLs
with minimal performance overhead, an idea which has come to be known as ``abstraction without regret''
\cite{lms,parreaux17,parreauxthesis, sheard99, trattdsl}.
This technique has been applied across a range of domains, including
Big data processing \cite{jet12}, stream processing \cite{moller20,strymonas},
query processing \cite{rhyme}, and parser combinators \cite{sspc, staged-parsers, krishnaswami19, flap}.
This work on staged parsers is the closest analogue to ours---indeed, PBT research has drawn an explicit link between parsing and generation, framing generators
as ``parsers of randomness'' \cite{goldsteinParsingRandomness2022}.
However, this equivalence is primarily theoretical. PBT generators are (a) not actually implemented as parsers
of random sequences, and (b) support choice-based combinators like \texttt{weighted\_union} that parsers do not.
For this reason, we must build staged generator libraries from scratch and cannot simply use existing staged parser libraries.
